---
title: "Equal Clusters"
author: "Nzambuli Daniel"
date: "2023-11-21"
output: html_document
---

# Data

```{r}
cluster_1 = c(1, 2, 1, 3, 3, 2, 1, 4, 1, 1)
cluster_2 = c(1, 3, 2, 2, 3, 1, 4, 1, 1, 2)
cluster_3 = c(2, 1, 1, 1,1, 3, 2, 1, 3, 1)
cluster_4 = c(1, 1, 3, 2, 1, 5, 1, 2, 3, 1)
data_1 = data.frame(clusters = c(rep("cluster_1", 10),
                               rep("cluster_2", 10),
                               rep("cluster_3", 10),
                               rep("cluster_4", 10)),
                  newspapers = c(cluster_1, cluster_2, cluster_3, cluster_4))
data_1
```

## Pivot the data 

```{r}
library(dplyr)
data_2 = data_1 %>% group_by(clusters) %>%
  summarise(mean = mean(newspapers),
            std = sd(newspapers))
data_2
```

## Mean of news papers in every household

```{r}
mean_house_newspaper = sum(data_2$mean)/ 4
mean_house_newspaper
```

`2` news papers are read in each house

## Anova analysis

use one sided anova to analyse the variance across the clusters.

```{r}
model = aov(newspapers~clusters, data = data_1)
summary(model)
```

## Variance estimate of population mean 

$$
var\ estimate\ mean = \hat{V}(\bar{y}_c) = (1 - \frac{n}{N}) * \frac{1}{n * m} * ms
$$

```{r}
ms = 0.3583
n = 4
N = 400
m = 10 # equal population size

var_eq_clust = (1 - n/N) * (1/ (n * m)) * ms
var_eq_clust
```

## Confidence Interval / Error Estimate 

$$
CI = 2 * \sqrt{\hat{v}(\bar{y}_c}
$$

```{r}
conf_int_eq_cluster = 2 * sqrt(var_eq_clust)
conf_int_eq_cluster
```

The confidence interval for the mean estimates for the total mean is $1.875 \pm 0.1883393$

## Standard Error

mean squared error is the residual for `mean_sq`

Therefore the mean squared error is `1.2028`

# Challenge 

suppose from a total of `20` bearing trees of guava in a village, `5 clusters` of size 4 trees each where selected and (hypothetical) yield (in kgs) is as given

```{r}
clust_1 = c(5, 4, 1, 15)
clust_2 = c(11, 1, 4, 7)
clust_3 = c(36, 10, 19, 11)
clust_4 = c(7, 15, 12, 10)
clust_5 = c(2, 22, 8, 6)

chall_data = data.frame(clusters = c(rep("cluster_1", 4),
                                     rep("cluster_2", 4),
                                     rep("cluster_3", 4),
                                     rep("cluster_4", 4),
                                     rep("cluster_5", 4)),
                        trees = c(clust_1, clust_2, clust_3, clust_4, clust_5))
head(chall_data)
```

### mean per cluster

```{r}
chall_data_2 = chall_data %>% group_by(clusters) %>%
  summarise(mean = mean(trees),
            std = sd(trees))
chall_data_2
```

```{r}
yield_per_tree = sum(chall_data_2$mean)/ nrow(chall_data_2)
yield_per_tree
```
